---
title: "Reproducible Research: Peer Assessment 1"
author: "GipFunny"
output: 
    html_document: 
        keep_md: true
---

<style>
hr { border-width: 3px; border-color: #333333; }
</style>

## Loading and preprocessing the data
```{r setup, include=FALSE}

# This section is not included in the output to make debugging easier.
# The next section has "eval=FALSE" to display the call to load data.

library(plyr)
library(dplyr)
library(tidyverse)
library(data.table)
library(lubridate)
library(ggplot2)

## Load Data
dfRaw <- read.csv(unzip("activity.zip"), header=TRUE, colClasses=c("integer", "Date", "integer"))
```

```{r loaddata, eval=FALSE}
## Load Data
dfRaw <- read.csv(unzip("activity.zip"), header=TRUE, colClasses=c("integer", "Date", "integer"))
```
***
## What is mean total number of steps taken per day?
```{r q1}
# remove all NA and make a copy
dtClean <- data.table(dfRaw[is.na(dfRaw$steps) == FALSE,])

# summarize by date
dtsums <- ddply(dtClean, ~date, summarize, sum=sum(steps))

# plot
ggplot(data=dtsums) + 
    geom_histogram(stat="identity", mapping=aes(x=date,y=sum)) +
    ggtitle("Total Steps Per Day") +
    ylab("Total Steps") + 
    xlab("Date")

# report mean and median of total steps per day
knitr::kable(
ddply(dtClean, ~date, summarize, mean=mean(steps), median=median(steps)), caption="Daily Mean and Median")
```

***
## What is the average daily activity pattern?
```{r q2}
# calculate mean by interval
dtInt <- ddply(dfRaw, ~interval, summarize, mean=mean(steps, na.rm=T))

plot(dtInt$interval, dtInt$mean, type="l", main="Average Daily Activity Pattern", xlab="Interval", ylab="Average Steps")

# max mean interval
maxMean <- dtInt[dtInt$mean == max(dtInt$mean),]
```
The interval `r maxMean$interval` has the maximum mean of `r maxMean$mean`.

***
## Imputing missing values

The total number of missing values is `r sum(is.na(dfRaw$steps))`.

``` {r q3}
# Make a copy
dfFull <- data.frame(dfRaw)

dfFuller <- cbind(dfFull, dtInt)
dfFuller$steps <- ifelse(is.na(dfFuller$steps), dfFuller$mean, dfFuller$steps)
dfFull <- dfFuller[, c(1,2,3)]

# summarize by date
dtsums2 <- ddply(dfFull, ~date, summarize, sum=sum(steps))

# plot
ggplot(data=dtsums2) + 
    geom_histogram(stat="identity", mapping=aes(x=date,y=sum)) +
    ggtitle("Total Steps Per Day") +
    ylab("Total Steps") + 
    xlab("Date")

# Get the maximum mean observation
maxMean2 <- dtInt[dtInt$mean == max(dtInt$mean),]

maxMeansAreEqual <- (maxMean$interval == maxMean2$interval & maxMean$mean == maxMean2$mean)

```

After imputing missing values the interval `r maxMean$interval` has the maximum average of `r maxMean$mean`.
The imputed maximum (impute NAs) and the previous maximum (ignore NAs) are `r ifelse(maxMeansAreEqual, "equal", "not equal")`.

***
## Are there differences in activity patterns between weekdays and weekends?

```{r q4}
# Weekdays
dfWeekDays <- data.frame(dfRaw)

# add weekday factor
dfWeekDays$day <- as.factor(ifelse(weekdays(dfWeekDays$date) %in% c("Saturday","Sunday"),"Weekend","Weekday"))

# calculate mean by interval and day
dtDayInt <- ddply(dfWeekDays, ~interval+day, summarize, mean=mean(steps, na.rm=T))

# plot with data and smooth line to highlight any pattern differences
ggplot(data=dtDayInt, mapping=aes(x=interval,y=mean)) + 
    geom_line() +
    facet_wrap(~day, nrow=2) +
    geom_smooth(se=F, method="loess", formula=y~x) +
    ggtitle("Total Steps Per Day") +
    ylab("Total Steps") + 
    xlab("Date")
```